\chapter{CONCLUSION}
In conclusion, a simple EEG-based BCI system was successfully designed and tested for basic typing control using blink and focus signals. The project showed that a basic BCI can be made even with a limited number of channels and simple hardware. The filter design was completed and tested to isolate the useful signal range, and signal processing methods such as amplification, filtering, and FFT analysis were used to study the EEG activity. The overall system performance was also tested using measures such as the confusion matrix and class performance, which helped show both the strengths and weaknesses of the system.

Although the system was able to work at a basic level, the results also showed several limitations that affected accuracy and reliability. These included the use of the same channel for blink and focus detection, the use of a matrix board instead of a PCB, fixed gain in the DRL stage, Arduino processing limitations, and distortion caused by non-ideal filter behavior. Even with these limitations, the project was useful because it gave practical understanding of how real biomedical signal systems behave outside of ideal theory.

This project was also a good learning experience in biomedical instrumentation. It helped build knowledge in areas such as biopotential signal acquisition, analog front-end circuit design, active filters, noise reduction, grounding, signal conditioning, and digital signal processing. Overall, the project showed how brain signals can be measured and used for simple control, while also teaching the challenges involved in building a real EEG-based system.